---
title: "Home_Work_3"
author: 'Group: Bob [Maryla Wozniak, Shakila Hoque, Elai Shalev, Matthew Perez]'
date: "10/5/2019"
output:
  pdf_document: default
  html_document: default
---
# GENERAL SETUP

```{r setup, include=FALSE}
## Turn off built in echo command for code chunks
## stops the 

# knitr::opts_chunk$set(echo = FALSE)

## Turn off the comments from the final output file
## This includes both the leading '#' from output 
## as well as comments in the code chuncks

# knitr::opts_chunk$set(comment = NA)
```

```{r, eval=FALSE}
## Clear all existing variables in global environment
rm(list = ls())
## Clear plot tab and close/save any open files 
dev.off()

## Step by step instructions
## Set the working dorectory to the location where the file was saved

## Setp: 1
## The 'rstudioapi' package allows you to access 
## system information using the R language
## the 'getSourceEditorContext' function allows you 
## to retreive informaton about the source files locaiton 
## this includes the file path without you having to
## know where the fiole is. THe computer doest that work for you.

## rstudioapi::getSourceEditorContext()


## Step: 2
## The line above is of type 'list' meaning there are multiple 
## peices of information contained within. In order to get the
## information we need we need to acess the file path 
## under the 'path' acesor in the return of the functioncall
## to do that we append '$path' this will print the file 
## path including the name of the file.

## rstudioapi::getSourceEditorContext()$path


## Step: 3
## In order to omit the name of the file from printing 
## we can use the 'dirname' function that will ignore 
## file names at the end of a file path and print
## only the path leading up to where the file is saved.

## dirname(rstudioapi::getSourceEditorContext()$path)


## Step: 4
## Lastly we use the 'setwd' comand to set the 
## directory to the returned value of the entire function
## call in order to set the working directory to the 
## locaiton of the file without ever having to look for where
## the file is actually stored.

## setwd(dirname(rstudioapi::getSourceEditorContext()$path))

## NOTE: THE LINES ABOVE ARE COMMENTED OUT IN ORDER TO 
## PREVENT COMPILATION ERRORS IN 'RMD' AND OTHER 'TEX' BASED 
## ENGINES. THIS IS A COMMAND IN ORDER TO SET YOUR WORKING 
## DIRECTORY AND SHOULD BE COMMENTED OUT OR COMPLETELY 
## OMITED FROM THE FINAL PRODUCT/PRESENTATION.
```

# IMPORT CLEAN DATA

Instead of importing and cleaning the data ourselves we are using the clean data contained in the ISLR packge from the CRAN website

```{r}
# Install the ISLR library so we can access the cleaned data set 
# for the following questions

# install.packages("ISLR")

# Save local copy of ISLR's Auto data  
Auto_local = ISLR::Auto

# Install and load "MASS" packge for aceess to the analytical functions later in the assignment
# install.packages("MASS")
library("MASS")

# Attach reference varables from data set 
attach(Auto_local)
```

\break
## Question 11: Develop a model to predict whether a car has high/low mpg.

#### (A) DEFINE BINARY VARIABLE 'mpg01' MPG01 = 1 IF MPG > MEDIAN(MPG) ELSE 0

```{r}
Auto_local[["mpg01"]] = as.integer(Auto_local$mpg > median(Auto_local$mpg))
summary(Auto_local)

```

#### (B) VISUAL DATA EXPLORATION
##### WHICH FEATURES ARE LIKELY TO HAVE THE GREATEST IMPACT? 
##### DESCRIBE YOUR FINDINGS

```{r}
## INITIAL PLOT
## The 'names' and 'mpg01' columns are not used in the following
## matrix plot because 'names' won't provide anymeaningful data in
## the visual analysis while 'mpg01' would be best analyzed using box plots 

pairs(Auto_local[,c(-9, -10)], lower.panel = NULL)

## COMMENTS:
## mpg01 is a direct product of mpg. 
## Since horsepower, weight, displacement, and acceleration 
## features affect mpg, they affect mpg01 as well. We boxplot 
## mpg and each of these features to investigate how they might affect mpg01


## COMMENTS:
## The boxplot below tells us that we can predict that cars with high 
## horsepower would have mpg01=0 (mpg below median). The highest 
## observation of the mpg01=1 category (neglecting outliers) is 
## about equal to the median observation of mpg01=0 category.
boxplot(horsepower~mpg01, 
        data = Auto_local, 
        ylab = "Horsepower",
        main = "Horsepower over mpg 01")


## COMMENTS:
## The boxplot below tells us that we can predict that heavy 
## cars would belong to the mpg01=0 category (mpg below median).
## The median observation of the mpg01=1 category is about 
## equal to the smallest observation of mpg01=0! If a car 
## weighs over 3500 we can certainly predict it will belong 
## to the mpg01=0 category, and half of the cars in the 
## Auto dataset weigh more than 3500. 
boxplot(weight~mpg01, 
        data = Auto_local, 
        ylab = "Weight", 
        main = "Weight over mpg 01")



## COMMENTS:
## The boxplot below tells us that we can predict that cars 
## with high displacement would belong to the mpg01=0 
## category (mpg below median). All of the observations 
## of the mpg01=1 category fit in the first quartile of 
## the mpg01=0 category. Using the displacement feature,
## we can predict over 75% of the cars would belong in 
## the mpg01=0 category. This feature seems to be the 
## most useful in trying to predict mpg01.
boxplot(displacement~mpg01, 
        data = Auto_local, 
        ylab = "Displacement",
        main = "Displacement over mpg 01")





## COMMENTS:
## The boxplot below tells us that it would be very hard 
## to use the acceleration feature to predict mpg01, 
## since the two categories distribute very similarly.

boxplot(acceleration~mpg01, 
        data = Auto_local, 
        ylab = "Acclereration",
        main = "Acclereration over mpg 01")
```

#### (C) SPLIT DATA INTO TEST AND TRAINING SETS

```{r}
## not true random but good enough 
set.seed(1)
num_train <- nrow(Auto_local) * 0.80

inTrain <- sample(nrow(Auto_local), size = num_train)

training <- Auto_local[inTrain,]
testing <- Auto_local[-inTrain,]

```

#### (D) PERFORM LDA ON THE TRAINGING SET TO PREDICT 'mpg01' 
###### WHAT IS THE TEST ERROR?

```{r}
## The following formula will be used for each 
## of the following models

fmla <- as.formula('mpg01 ~ horsepower + weight + displacement')

lda_model <- lda(fmla, data = training)

lda_pred <- predict(lda_model, testing)

table(lda_pred$class, testing$mpg01)

1 - mean(lda_pred$class == testing$mpg01)

## COMMENTS:
#The LDA model works great!
#The test error is 0.1265823

```

#### (E) PERFORM QDA ON THE TRAINGING SET TO PREDICT 'mpg01' 
###### WHAT IS THE TEST ERROR?

```{r}
qda_model = qda(fmla, data = training)

qda_pred <- predict(qda_model, testing)

table(qda_pred$class, testing$mpg01)

1 - mean(qda_pred$class == testing$mpg01)

## COMMENTS:
## QDA modal test error is lower then the LDA, 
## this is likely because QDA is a quadratic line
## meaning the raltionship between the variables 
## that we ploted dont have a linear relationship.
## The error rate is 0.1012658


```

#### (F) PERFORM LOGISTIC REGRESSION ON THE TRAINING DATA IN ORDER TO PREDICT 'mpg01'
###### WHAT IS THE TEST ERROR?

```{r}
log_reg <- glm(fmla, data = training, family = binomial)

pred <- predict(log_reg, testing, type = 'response')
pred_values <- round(pred)
table(pred_values, testing$mpg01)

1- mean(pred_values == testing$mpg01)

## COMMENTS:
## The Legistic Regression model performed 
## considerably better then both the LDA
## and the QDA models. Our asssumption is that 
## because the Logistic regression is designed to 
## handle non-numeric data it's better suited to conduct 
## the analysis for this exercise.
## The error rate is 0.06329114

```