---
title: "Home_Work_3"
author: 'Group: Bob [Maryla Wozniak, Shakila Hoque, Elai Shalev, Matthew Perez]'
date: "10/5/2019"
output: html_document
---
# GENERAL SETUP

```{r}
rm(list = ls())
dev.off()
```

# IMPORT CLEAN DATA

First we need to install some packages

```{r}
# install.packages("MASS"")
# install.packages("ISLR")
library("MASS")
library("ISLR")
Auto_local = Auto
attach(Auto_local)
```

\break
# Question 11: Develop a model to predict whether a car has high/low mpg.
## (A) DEFINE BINARY VARIABLE 'mpg01' MPG01 = 1 IF MPG > MEDIAN(MPG) ELSE 0

```{r}
mpg01 = factor(mpg, ordered = F, levels = c(0,1))
mpg01 = ifelse(mpg>median(mpg), 1, 0)
Auto_local = data.frame(Auto_local,mpg01)
summary(Auto_local)
```

## (B) VISUAL DATA EXPLORATION
### WHICH FEATURES ARE LIKELY TO HAVE THE GREATEST IMPACT? 
### DESCRIBE YOUR FINDINGS

```{r}
pairs(Auto_local, lower.panel = NULL)

## mpg01 is a direct product of mpg. 
## Since horsepower, weight, displacement, and acceleration features affect mpg, they affect mpg01 as well.
## We boxplot mpg and each of these features to investigate how they might affect mpg01

boxplot(horsepower~mpg01, data = Auto_local, main = "Horsepower and mpg Boxplot")

## above boxplot tells us that we can predict that cars with high horsepower would have mpg01=0 (mpg below median).
## the highest observation of the mpg01=1 category (neglecting outliers) is about equal to the median observation of mpg01=0 category.

boxplot(weight~mpg01, data = Auto_local, main = "Weight and mpg Boxplot")

## above boxplot tells us that we can predict that heavy cars would belong to the mpg01=0 category (mpg below median).
## the median observation of the mpg01=1 category is about equal to the smallest observation of mpg01=0! 
## If a car weighs over 3500 we can certainly predict it will belong to the mpg01=0 category, and half of the cars in the Auto dataset weigh more than 3500. 

boxplot(displacement~mpg01, data = Auto_local, main = "Displacement and mpg Boxplot")

## above boxplot tells us that we can predict that cars with high displacement would belong to the mpg01=0 category (mpg below median).
## all of the observations of the mpg01=1 category fit in the first quartile of the mpg01=0 category.
## Using the displacement feature, we can predict over 75% of the cars would belong in the mpg01=0 category.
## This feature seems to be the most useful in trying to predict mpg01.

boxplot(acceleration~mpg01, data = Auto_local, main = "Acclereration and mpg Boxplot")

## above boxplot tells us that it would be very hard to use the acceleration feature to predict mpg01, since the two categories distribute very similarly.








```

## (C) SPLIT DATA INTO TEST AND TRAINING SETS

```{r}
set.seed(1)
num_train <- nrow(Auto_local) * 0.75

inTrain <- sample(nrow(Auto_local), size = num_train)

training <- Auto_local[inTrain,]
testing <- Auto_local[-inTrain,]

training
testing

# First, we are omitting the cars with 3 or 5 cylinders becasue there are only few of them, 
#it will likely skew the analysis. 

```

## (D) PERFORM LDA ON THE TRAINGING SET TO PREDICT 'mpg01' 
### WHAT IS THE TEST ERROR?

```{r}
require(MASS)
fmla <- as.formula('mpg01 ~ displacement + horsepower + weight + year + cylinders')
lda_model <- lda(fmla, data = training)

lda_pred <- predict(lda_model, testing)

table(lda_pred$class, testing$mpg01)

1 - mean(lda_pred$class == testing$mpg01)

#The LDA model works great!
#The test error is 0.122449 

```

## (E) PERFORM QDA ON THE TRAINGING SET TO PREDICT 'mpg01' 
### WHAT IS THE TEST ERROR?

```{r}
qda_model = qda(fmla, data = training)

qda_pred <- predict(qda_model, testing)

table(qda_pred$class, testing$mpg01)

1 - mean(qda_pred$class == testing$mpg01)

# QDA modal test error is still the same as LDA, however the test results for each value is different.
# Example; the LDA model has a higher rate of error for values over median but the QDA have more equal distribution 
#in the error for test value. 


```

## (F) PERFORM LOGISTIC REGRESSION ON THE TRAINING DATA IN ORDER TO PREDICT 'mpg01'
### WHAT IS THE TEST ERROR?

```{r}
log_reg <- glm(fmla, data = training, family = binomial)

pred <- predict(log_reg, testing, type = 'response')
pred_values <- round(pred)
table(pred_values, testing$mpg01)

1- mean(pred_values == testing$mpg01)

#The Legistic Regressional model seems similar to LDA and QDA model, eventhough it did not perform as well as other two 
#model. The error rate is 0.09183673


```

# OPTIONAL
## (G) PERFORM KNN ON THE TRAINING DATA WITH SEVERAL VALUES FOR 'K' IN ORDER TO PREDICS 'mpg01' 
### WHAT IS THE TEST ERROR?
### WHAT VALUE OF K PERFORMS BEST?

```{r}

```
