---
title: "Home_Work_3"
author: 'Group: Bob [Maryla Wozniak, Shakila Hoque, Elai Shalev, Matthew Perez]'
date: "10/5/2019"
output: html_document
---
# GENERAL SETUP

```{r}
rm(list = ls())
dev.off()
```

# IMPORT CLEAN DATA

First we need to install some packages

```{r}
# install.packages("MASS"")
# install.packages("ISLR")
library("MASS")
library("ISLR")
Auto_local = Auto
attach(Auto_local)
```

\break
# Question 11: Develop a model to predict whether a car has high/low mpg.
## (A) DEFINE BINARY VARABLE 'mpg01' MPG01 = 1 IF MPG > MEDIAN(MPG) ELSE 0

```{r}
Auto_local[["mpg01"]] =  as.integer(Auto_local$mpg > median(Auto_local$mpg))
```

## (B) VISUAL DATA EXPLORATION
### WHICH FEATURES ARE LIKELY TO HAVE THE GREATEST IMPACT? 
### DESCRIBE YOUR FINDINGS

```{r}
# PAIRWISE MATRIX PLOT EXCLUDING MPG AND NAME
pairs(Auto_local[, c(-1, -2,-8,-9, -10)], lower.panel = NULL)


```

## (C) SPLIT DATA INTO TEST AND TRAINING SETS

```{r}

```

## (D) PERFORM LDA ON THE TRAINGING SET TO PREDICT 'mpg01' 
### WHAT IS THE TEST ERROR?

```{r}

```

## (E) PERFORM QDA ON THE TRAINGING SET TO PREDICT 'mpg01' 
### WHAT IS THE TEST ERROR?

```{r}

```

## (F) PERFORM LOGISTIC REGRESSION ON THE TRAINING DATA IN ORDER TO PREDICT 'mpg01'
### WHAT IS THE TEST ERROR?

```{r}

```

# OPTIONAL
## (G) PERFORM KNN ON THE TRAINING DATA WITH SEVERAL VALUES FOR 'K' IN ORDER TO PREDICS 'mpg01' 
### WHAT IS THE TEST ERROR?
### WHAT VALUE OF K PERFORMS BEST?

```{r}

```
